Ornaments as indicators of social changes and cultural practices in northeastern Taiwan before and after European colonial period

### 2019-12-10 analysis/paper/paper.Rmd
some more edits

Kiwulan (Figure \@ref(fig:KWL-map)) was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge. The excavation includes eight open areas in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007]. The nearly 2 m thick archaeological deposits reveal a large amount of artefacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-tern settlement. The center of the site is the open area consists of A and D section which is also the study area where our samples selected from. In the AD area, post-holes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures. At the north margin of the dwelling place were burials that are mostly oriented in an east-west direction (Figure \@ref(fig:KWL-sam-area)).

@@

The archaeological indicators of the start of the European influence at Kiwulan are the appearance of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century. A large amount of light gray glazed jars were found on the Spanish shipwreck San Diego sinking in 1600 AD, which is believed to be the earliest evidence yet found [@Hsieh2009; @Hsieh1995]. The origin of light gray glazed jars was thought from Southeast China and commonly found at sites in Taiwan that were associated with the European activities, such as Zeelandia fort site in Tainan [@Wang2007]. Regarding large dark brown glazed stoneware jars, they might be made in Southeast Asia and are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages. The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131]. They might be directly imported to Kiwulan by the Europeans to Kiwulan or indirectly by Chinese merchants or other indigenous groups via the regional network in north Taiwan. Either way indicates that the Europeans played an important role in introducing foreign jars to Taiwan. Those jars were widely distributed across the site and served as indicators, together with the radiocarbon dates, to identify the excavation unites associated with the pre-European and the start of the European influence at Kiwulan. The archaeological signature of the Chinese period at Kiwulan is the large amount and diversity of Chinese porcelains in many forms such as bowls, plates, and cups. Other indicators include opium pipe-bowls and distinctive architectural bricks and tiles used by Chinese [@Hsieh2009]. Chinese migrations to Yilan were also recorded in official Chinese records written in the early 19th century recording the first immigrant in 1768 [@Chen1963; @Ke1993].

@@

The ornaments found in the Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament are golden beads (n = `r golden_bead_n`), followed by glass beads (n =  `r glass_bead_n`), carnelian beads (n = `r agate_bead_n`), metal rings (n = `r metal_ring_n`), and metal bells (n = `r metal_bell_n`, see also Figure \@ref(fig:ornaments-image)). The variety of subtypes and their individual frequencies are summarized in Table \@ref(tab:tidy-data-for-table). The criteria for classification of subtype is based on their dimensional attributes that are important to understand the variation in shapes. This might simplify the subtypes of glass beads when they are the same size and shape but in different color or texture. However, their variety would be considered and discussed separately. The ornaments found in the dwelling area were assumed as body adornments or a part of clothing in their daily life according to their use in the burial contexts.

@@

An indirect colonial influence is indicated at Kiwulan by a greater diversity of ornament types and materials during the European period. Northeastern Taiwan, Yilan, was involved in complex trading networks both in regional scale with other indigenous groups and Chinese merchants, and global scale with the European including the Dutch and the Spanish. The origin of ornaments might be multiple places mostly in Southeast Asia and introduced into northeastern Taiwan by Chinese merchants before 17th century. Later, the trade activities became more frequent and intense in the 17th century that might be related to the colonial involvement of the Europeans based on the similar styles of golden beads found in Luzon that was once the colony of the Spanish during this period. In addition to ornaments, stonewares at Kiwulan that were also found on a Spanish shipwreck and Southeast Asia in relation to the Dutch and the Spanish trading routes, and the tobacco use were also believed introduced to Kiwulan by the Europeans [@Wang2007; @Chen2007]. The greater diversity and quantity of ornaments likely results from participation in large scale exchange networks that stimulated the circulation of different ornament classes. However, the frequency of overall ornaments and each subtype declines significantly during the Chinese period in the early 19th century. This may be due to a smaller scale of trading network, the overall decline of the indigenous population in Yilan, and adopting the practices of Han Chinese who did not use these kinds of beads in their dress. Since the end of the 18th century, many indigenous people moved southwards to Hualien due to population pressure caused by Han Chinese immigrants [@Chen2007]. Although historical records indicate that trade ornaments were prestige goods in prehistoric northeastern Taiwan, the information might be biased that needs to be compared with archaeological evidence.

@@

Another scenario could be that the increasing number and concentrated spatial pattern of ornaments indicate cultural resistance against the European intrusion by using them to show social identity and emphasize the local custom that had existed before European contact. Resistance could be presented in many forms. For example, @Rubertone2000 discussed a mid-17th century Narragansett Indian cemetery where a large amount of native-made shell beads supposed to be tributes for the European were found in the graves as an act of political resistance against the colonial demands. Similarly, the frequent presence of ornaments across the dwelling area during the European suggests ornaments might be intentionally displayed in daily life to reinforce the original cultural tradition. During the Chinese period in the 19th century, the decreasing frequency and diversity of ornaments show a decline in the production, use and discard of ornaments at Kiwulan. This might reflect the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals to accumulate wealth, or population decrease. It is also possible that some ornaments were treated as heirlooms, such as carnelian beads and golden beads, that passed from one generation to the next. This might suggest that the ornaments found in the Chinese period were old items continuously used until then. Further analysis of the source for ornaments from different periods will be helpful to understand their origin and regional circulation.

>>>>

Kiwulan (Figure \@ref(fig:KWL-map)) was a rescue archaeology project that carried out from 2001 to 2004 in advance of water diversion project and construction of a road bridge. The excavation includes eight open area sections in total of 262 squares (4 m by 4 m) reaching 3,814 m^2^ [@Chen2007]. The nearly 2 m thick archaeological deposits reveal a large amount of artefacts, burials, middens, post-holes, wooden pillars, and stone structures, all of which indicates it was a long-tern settlement. Based on the continuous deposition and the frequency of artifacts, the center of the site is the open area consisting of A and D sections, which is also the study area where our samples selected from. In the AD area, post-holes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures. At the north margin of the dwelling place were burials that are mostly oriented in an east-west direction (Figure \@ref(fig:KWL-sam-area)).

@@

The archaeological indicators of the start of the European influence at Kiwulan are the appearance of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century. A large amount of light gray glazed jars were found on the Spanish shipwreck San Diego sinking in 1600 AD, which is believed to be the earliest evidence yet found [@Hsieh2009; @Hsieh1995]. The origin of light gray glazed jars was thought from Southeast China and commonly found at sites in Taiwan that were associated with the European activities, such as Zeelandia fort site in Tainan [@Wang2007]. Regarding large dark brown glazed stoneware jars, they might be made in Southeast Asia and are frequently found in European shipwrecks from this period for transporting water, wine or other fluids on the long voyages. The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131]. They might be directly imported to Kiwulan by the Europeans to Kiwulan or indirectly by Chinese merchants or other indigenous groups via the regional network in north Taiwan. Either way indicates that the Europeans played an important role in introducing foreign jars to Taiwan. Those jars were widely distributed across the site and served as indicators, together with the radiocarbon dates, to identify the excavation unites associated with the pre-European and the start of the European influence at Kiwulan. In addition to ornaments and stoneware jars, around 300 pieces of local made pipes and a few imported pipes were viewed as a custom introduced by the Europeans.

The archaeological signature of the Chinese period at Kiwulan is the large amount and diversity of Chinese porcelains in many forms such as bowls, plates, and cups. Other indicators include opium pipe-bowls and distinctive architectural bricks and tiles used by Chinese [@Hsieh2009]. Chinese migrations to Yilan were also recorded in official Chinese records written in the early 19th century recording the first immigrant in 1768 [@Chen1963; @Ke1993].

@@

The ornaments found in the Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament are golden beads (n = `r golden_bead_n`), followed by glass beads (n =  `r glass_bead_n`), carnelian beads (n = `r agate_bead_n`), metal rings (n = `r metal_ring_n`), and metal bells (n = `r metal_bell_n`, see also Figure \@ref(fig:ornaments-image)). The variety of subtypes and their individual frequencies are summarized in Table \@ref(tab:tidy-data-for-table). The criteria for classification of subtype is based on their dimensional attributes that are important to understand the variation in shapes. This might simplify the subtypes of glass beads when they are the same size and shape but in different color or texture. However, the variation would be considered and discussed separately. The ornaments found in the dwelling area were assumed as body adornments or a part of clothing in their daily life according to the use in burial contexts.

@@

An indirect colonial influence is indicated at Kiwulan by a greater diversity of ornament types and materials during the European period. Northeastern Taiwan, Yilan, was involved in complex trading networks both in regional scale with other indigenous groups and Chinese merchants, and global scale with the European including the Dutch and the Spanish. The origin of ornaments might be multiple places mostly in Southeast Asia and introduced into northeastern Taiwan by Chinese merchants before 17th century. Later, the trade activities became more frequent and intense in the 17th century that might be related to the colonial involvement of the Europeans based on the similar styles of golden beads found in Luzon that was once the colony of the Spanish during this period. In addition to ornaments, stonewares at Kiwulan that were also found on a Spanish shipwreck and Southeast Asia in relation to the Dutch and the Spanish trading routes, and the tobacco use were also believed introduced to Kiwulan by the Europeans [@Wang2007; @Chen2007]. The greater diversity and quantity of ornaments likely results from participation in large scale exchange networks that stimulated the circulation of different ornament classes. However, the frequency of overall ornaments and each subtype declines significantly during the Chinese period in the early 19th century. This may be due to a smaller scale of trading network, the overall decline of the indigenous population in Yilan, or adopting the practices of Han Chinese who did not use these kinds of beads in their dress. Since the end of the 18th century, many indigenous people moved southwards to Hualien due to population pressure caused by Han Chinese immigrants [@Chen2007]. Although historical records indicate that trade ornaments were prestige goods in prehistoric northeastern Taiwan, the information might be biased that needs to be compared with archaeological evidence.

@@

Another scenario could be that the increasing number and concentrated spatial pattern of ornaments indicate cultural resistance against the European intrusion by using them to show social identity and emphasize the local custom that had existed before European contact. Resistance could be presented in many forms. For example, @Rubertone2000 discussed a mid-17th century Narragansett Indian cemetery where a large amount of native-made shell beads supposed to be tributes for the European were found in the graves as an act of political resistance against the colonial demands. Similarly, the frequent presence of ornaments across the dwelling area during the European period suggests ornaments might be intentionally displayed in daily life to reinforce the original cultural tradition. During the Chinese period in the 19th century, the decreasing frequency and diversity of ornaments show a decline in the use and discard of ornaments at Kiwulan. This might reflect the smaller scale of exchange networks, with limited sources of ornaments, and limited opportunities for individuals to accumulate wealth, or indeigenous population decline. It is also possible that some ornaments were treated as heirlooms, such as carnelian beads and golden beads, that passed from one generation to the next. This might suggest that the ornaments found in the Chinese period were old items continuously used until then. Further analysis of the source for ornaments from different periods will be helpful to understand their origin and regional circulation.
